# Reproducible Research: Peer Assessment 1

```{r setoptions,echo=FALSE}
opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data

```{r}
# create the data folder

if (!file.exists("data")) {
        dir.create("data")
        }

# unzip the file use utils package use "install.packages('utils')" to install

library(utils)
unzip('./activity.zip', exdir = './data/')

# read the data

DF <- read.csv('./data/activity.csv')

# change the class of 'date' variable from factor to date

DF$date <- as.Date(DF$date,'%Y-%m-%d')
```

I loaded the data into a data frame called DF with the variable 'date' changed to
a date class.

## What is mean total number of steps taken per day?

```{r}
# ignore the NA

DFn <- DF[!is.na(DF$steps),]

# make a histogram of the total number of steps taken each day

library(lattice)
with(DFn,histogram(~steps|date))
```

Mean total number of steps taken per day could be get from the following code.

```{r}
# calculate and report the mean and median total number of steps taken per day

with(DFn,tapply(steps,date,mean))

with(DFn,tapply(steps,date,median))

```

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)

```{r}
plot(unique(DFn$interval),with(DFn,tapply(steps,interval,mean)),
     xlab='Interval',
     ylab='Numbers of steps',
     type='l')
```

Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?

```{r}
names(which.max(with(DFn,tapply(steps,interval,mean))))
```

It seems that around 8:35AM is the most active time segment for the anonymous 
individual.

## Imputing missing values

Note that there are a number of days/intervals where there are missing values
(coded as NA). The presence of missing days may introduce bias into some
calculations or summaries of the data.

The total number of missing values in the dataset is get by the following code.
In fact, only the steps have NA.

```{r}
sum(!complete.cases(DF))
```

Use the mean for that 5-minute interval to fill in the missing NA for there are 
patterns during a day and create a new dataset called DFnew with the missing 
data filled in.

```{r}
group <- split(DF,DF$interval)
DFnew <- data.frame(DF[0,])
findNA <- function (x){
        x[which(!complete.cases(x)),1] <- mean(x[-which(!complete.cases(x)),1])
        x
}
for (i in 1:288){
        DFnew <- rbind(DFnew,findNA(group[[i]]))
}
```

Plot histogram of the total number of steps taken each day and calculate the 
mean and median total number of steps taken per day. 

```{r}
library(lattice)
with(DFnew,histogram(~steps|date))
with(DFnew,tapply(steps,date,mean))
with(DFnew,tapply(steps,date,median))
```

These values are differ from the estimates from the first part of the assignment
depending on the impute method.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
# add 'weekday' variable

DFnew$weekday <- 'weekend'
DFnew[!(weekdays(DFnew$date)=="Sunday"|weekdays(DFnew$date)=="Saturday"),]$weekday <- 'weekday'

# plot the differences and the log 10 based version in the comment line

library(lattice)
with(DFnew,xyplot(steps~interval|weekday,type='l',layout=c(1,2),ylab='Numbers of steps'))
# with(DFnew,xyplot(steps~interval|weekday,type='l',layout=c(1,2),ylab='Numbers of steps'))
```

The differeces is clear from the plot, the steps is earlier on weekday and the
anonymous individual may like to walk on weekend evening. I noticed the figure 
in the instruction use a log based transform which I don't think is useful for 
this problem to find the patterns.

